---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rpigpior

<!-- badges: start -->

<!-- badges: end -->

`rpigpior` allows the R programming language access to the Raspberry Pi
GPIO using libgpiod. Because libgpiod is already included in Raspbian,
there is nothing else to install.

rpigpior implements the six tools provided by
[libgpiod](https://git.kernel.org/pub/scm/libs/libgpiod/libgpiod.git/about/)
for GPIO access:

-   **gpiodetect** - list all gpiochips present on the system, their
    names, labels and number of GPIO lines

-   **gpioinfo** - list lines, their gpiochip, offset, name, and
    direction, and if in use then the consumer name and any other
    configured attributes, such as active state, bias, drive, edge
    detection and debounce period

-   **gpioget** - read values of specified GPIO lines

-   **gpioset** - set values of specified GPIO lines, holding the lines
    until the process is killed or otherwise exits

-   **gpiomon** - wait for edge events on GPIO lines, specify which
    edges to watch for, how many events to process before exiting, or if
    the events should be reported to the console

-   **gpionotify** - wait for changed to the info for GPIO lines,
    specify which changes to watch for, how many events to process
    before exiting, or if the events should be reported to the console

## Installation

`rpigpior` is hosted on github. Until it arrives at cran, you can
install the development version of rpigpior from
[GitHub](https://github.com/mnr/rpigpior) with:

``` r
remotes::install_github("mnr/rpigpior")
```

Most documentation will tell you to use devtools, but
`install.packages("devtools")` is an onerous task.

## Example

### gpiodetect

```{r}
gpiodetect()
```
