---
title: "Cookbook"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rpigpior)
```

This cookbook shows how to use sensors and drivers using tools provided by `rpigpior`.

# BME280

`rpi_i2c_get(chip_address = 0x77, data_address= 0xD0)` returns the BME Chip ID, 0x60.

#rpi_i2c_get(chip_address = 0x77, data_address = 0xD0) \# read chip id control_hum_address \<- 0xf2 sample_hum \<- 0x01 control_measure_address \<- 0xf4 control_press_temp \<- strtoi("00100101", base = 2)

# setup humidity measurement

rpi_i2c_set(chip_address = 0x77, data_address = control_hum_address, value = sample_hum, data_size = "b")

# setup and execute pressure & temperature measurement

rpi_i2c_set(chip_address = 0x77, data_address = control_measure_address, value = control_press_temp, data_size = "b")

result_pressure \<-rpi_i2c_get(chip_address = 0x77, data_address = 0xf7, data_size = "w")

result_temperature \<-rpi_i2c_get(chip_address = 0x77, data_address = 0xfa, data_size = "w")

result_humidity \<-rpi_i2c_get(chip_address = 0x77, data_address = 0xfd, data_size = "w")

print(paste("raw pressure:", as.numeric(result_pressure))) print(paste("raw temperature:", as.numeric(result_temperature))) print(paste("raw humidity:", as.numeric(result_humidity)))

# convert raw temp to centigrade

\# Calibration coefficients

dig_T1 \<- 27384

dig_T2 \<- 25740

\# Raw temperature data (16-bit signed integer)

raw_temperature \<- 519888

\# Temperature compensation

var1 \<- (raw_temperature / 16384 - dig_T1 / 1024) \* dig_T2

var2 \<- ((raw_temperature / 131072 - dig_T1 / 8192)\^2) \* dig_T2

t_fine \<- var1 + var2

\# Calculate temperature in degrees Celsius

temperature_celsius \<- (t_fine / 5120)

\# Print the result

cat("Temperature in Celsius:", temperature_celsius, "°C\\n")

# convert raw pressure to Pa

\# Calibration coefficients

dig_P1 \<- 36477

dig_P2 \<- -10476

dig_P3 \<- 3024

dig_P4 \<- 2855

dig_P5 \<- 140

dig_P6 \<- -7

dig_P7 \<- 15500

dig_P8 \<- -14600

dig_P9 \<- 6000

\# Raw pressure data (20-bit unsigned integer)

raw_pressure \<- 416231

\# Temperature compensation (from previous calculation)

t_fine \<- 128422

\# Calculate pressure

var1 \<- t_fine / 2 - 64000

var2 \<- var1\^2 \* dig_P6 / 32768

var2 \<- var2 + (var1 \* dig_P5) \* 2

var2 \<- (var2 / 4) + (dig_P4 \* 65536)

var1 \<- (dig_P3 \* var1\^2 / 524288 + dig_P2 \* var1) / 524288

var1 \<- (1 + var1 / 32768) \* dig_P1

pressure := 1048576 - raw_pressure

pressure \<- ((pressure - (var2 / 4096)) \* 6250) / var1

\# Print the result

cat("Pressure in Pa:", pressure, "Pa\\n")

# humidity to relative humidity

\# Calibration coefficients

dig_H1 \<- 75

dig_H2 \<- 347

dig_H3 \<- 0

dig_H4 \<- 320

dig_H5 \<- 0

dig_H6 \<- 30

\# Raw humidity data (16-bit unsigned integer)

raw_humidity \<- 4162

\# Temperature compensation (from previous calculation)

t_fine \<- 128422

\# Calculate humidity

temp_compensated_humidity \<- (t_fine - 76800)

temp_compensated_humidity \<- (raw_humidity - (dig_H4 \* 64 + dig_H5 / 16384 \* temp_compensated_humidity)) \* (dig_H2 / 65536 \* (1 + dig_H6 / 67108864 \* temp_compensated_humidity \* (1 + dig_H3 / 67108864 \* temp_compensated_humidity)))

temp_compensated_humidity \<- temp_compensated_humidity \* (1 - dig_H1 \* temp_compensated_humidity / 524288)

humidity_percentage \<- ifelse(temp_compensated_humidity \> 100, 100, temp_compensated_humidity)

humidity_percentage \<- ifelse(humidity_percentage \< 0, 0, humidity_percentage)

\# Print the result

cat("Relative Humidity:", humidity_percentage, "%\\n")
