---
title: "Raspberry Pi and PWM"
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rpigpior)
```

# Pulse Width Modulation on the Raspberry Pi

## Hardware PWM vs Software PWM

The Raspberry Pi exposes two PWM channels from the Broadcom chip. These
appear on lines which also supply GPIO connections and so the Broadcom
ALT functions need to be called.

PWM0 appears on board pins 12 (GPIO18) and 32 (GPIO12). PWM1 appears on
board pins 33 (GPIO13) and 35 (GPIO19). This causes a confusing
situation in two ways:

-   Don't confuse board pin 12 with GPIO12 (board pin 32). Both are
    connected to BCM PWM0

The Raspberry Pi analog audio output uses both PWM channels. You can't
use hardware PWM and audio output at the same time.

Although it is possible to create software PWM, this isn't advised due
to latency of linux and R.

`rpi_pwm()` handles all of these issues. If you stick to board pin
numbers, it will all work out. If you go astray, `rpi_pwm()` will `stop`
and warn you.

## Enabling hardware PWM on the Raspberry Pi

edit /boot/config.txt as root

`sudo nano /boot/config.txt`

`rpi_pwm(pin_number = 12, mode = "enable")`

[Enabling one PWM
channel](https://github.com/raspberrypi/linux/blob/04c8e47067d4873c584395e5cb260b4f170a99ea/arch/arm/boot/dts/overlays/README#L925)

[Enabling a dual pwm
channel](https://github.com/raspberrypi/linux/blob/04c8e47067d4873c584395e5cb260b4f170a99ea/arch/arm/boot/dts/overlays/README#L944)

If you need more details on the PWM channels on the Raspberry Pi, [here
is an excellent
write-up](https://github.com/dotnet/iot/blob/main/Documentation/raspi-pwm.md).

## How to use `rpi_pwm()`

<https://www.kernel.org/doc/html/v5.10/driver-api/pwm.html> \<\> GPIO18
(pin 12) and GPIO12 (pin 32) = PWM0 GPIO13 (pin 33) and GPIO19 (pin 35)
= PWM1

Only two PWM lines are available on the Raspberry Pi - pwm0 and pwm1.

Select one or two of four hardware PWM lines (12, 32, 33, or 35).
Generates an error message if a pwm channel is selected twice. \#' Since
PWM0 appears on both pins 12 and 32, and PWM1 appears on pins 33 and 35,
a valid selection of two pins is one from PWM0 and one from PWM1. \#' In
other words, only these combinations are valid: \#' \itemize{
#'        \item pin_number = c(12,33)
#'        \item pin_number = c(32,33)
#'        \item pin_number = c(12,35)
#'        \item pin_number = c(32,35)
#'     }
