---
title: "Raspberry Pi and PWM"
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rpigpior)
```

# Pulse Width Modulation on the Raspberry Pi

## Hardware PWM vs Software PWM

The Raspberry Pi exposes two PWM channels from the Broadcom chip. These
appear on lines which also supply GPIO connections and so the Broadcom
ALT functions need to be called.

-   PWM0 appears on board pins 12 (GPIO18) and 32 (GPIO12).

-   PWM1 appears on board pins 33 (GPIO13) and 35 (GPIO19).

This causes a confusing situation in two ways:

-   Don't confuse board pin 12 with GPIO12 (board pin 32). Both are
    connected to BCM PWM0
-   If you want to use two PWM channels, there are only certain
    combinations of pins that will provide the two channels. These
    combinations are board pins (12,33), (32,33), (12,35), or (32,35)

`rpi_pwm()` handles these issues. If you stick to board pin numbers, it
will all work out. If you go astray, `rpi_pwm()` will `stop` and warn
you.

The Raspberry Pi analog audio output uses both PWM channels. You can't
use hardware PWM and audio output at the same time.

Although it is possible to create software PWM, this isn't advised due
to latency of linux and R. `rpigpior` does not supply any software PWM
channels.

## Enabling hardware PWM on the Raspberry Pi

*Raspberry Pi OS* uses the device tree to enable hardware PWM. You'll
need to do a bit of behind-the-scenes editing to set this up.

To do this, edit /boot/config.txt as root.

1.  On the Raspberry Pi, open a terminal window

2.  Use the command `sudo nano /boot/config.txt`. This will open the
    nano text editor and display the contents of /boot/config.txt.

3.  Past the `dtoverlay` command at the end of the file. The easiest way
    to determine the command to add to `/boot/config.txt` is to ask
    `rpi_pwm()` for help. Select the pins you want to use and if they
    aren't enabled, `rpi_pwm()` will stop and provide you with the
    correct `dtoverlay` string.

4.  Use *control-o* to write the file, then *control-x* to exit `nano`

If you're curious about how or why this `dtoverlay` is built, take a
look at this excellent reference.

[Enabling one PWM
channel](https://github.com/raspberrypi/linux/blob/04c8e47067d4873c584395e5cb260b4f170a99ea/arch/arm/boot/dts/overlays/README#L925)

[Enabling a dual pwm
channel](https://github.com/raspberrypi/linux/blob/04c8e47067d4873c584395e5cb260b4f170a99ea/arch/arm/boot/dts/overlays/README#L944)

If you need more details on the PWM channels on the Raspberry Pi, [here
is an excellent
write-up](https://github.com/dotnet/iot/blob/main/Documentation/raspi-pwm.md).

## How to use `rpi_pwm()`

<https://www.kernel.org/doc/html/v5.10/driver-api/pwm.html> \<\> GPIO18
(pin 12) and GPIO12 (pin 32) = PWM0 GPIO13 (pin 33) and GPIO19 (pin 35)
= PWM1

Only two PWM lines are available on the Raspberry Pi - pwm0 and pwm1.

Select one or two of four hardware PWM lines (12, 32, 33, or 35).
Generates an error message if a pwm channel is selected twice. \#' Since
PWM0 appears on both pins 12 and 32, and PWM1 appears on pins 33 and 35,
a valid selection of two pins is one from PWM0 and one from PWM1. \#' In
other words, only these combinations are valid: \#' \itemize{
#'        \item pin_number = c(12,33)
#'        \item pin_number = c(32,33)
#'        \item pin_number = c(12,35)
#'        \item pin_number = c(32,35)
#'     }

## Handling Errors:

### Invalid PWM pin:

### Invalid PWM pin combination:

### PWM not enabled:

The PWM channel (or channels) you have selected are not yet enabled.
Enabling PWM is described in the above section titled [Enabling hardware
PWM on the Raspberry Pi] . The error message supplies the string you'll
need to paste into `/boot/config.txt`.
