---
title: "Raspberry Pi and PWM"
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rpigpior)
```

# Pulse Width Modulation on the Raspberry Pi

## Hardware PWM vs Software PWM

The Raspberry Pi exposes two PWM channels from the Broadcom chip. These
appear on lines which also supply GPIO connections and so the Broadcom
ALT functions need to be called.

-   PWM0 appears on board pins 12 (GPIO18) and 32 (GPIO12).

-   PWM1 appears on board pins 33 (GPIO13) and 35 (GPIO19).

This causes a confusing situation in two ways:

-   Don't confuse board pin 12 with GPIO12 (board pin 32). Both are
    connected to BCM PWM0
-   If you want to use two PWM channels, there are only certain
    combinations of pins that will provide the two channels. These
    combinations are board pins (12,33), (32,33), (12,35), or (32,35)

`rpi_pwm()` handles these issues. If you stick to board pin numbers, it
will all work out. If you go astray, `rpi_pwm()` will `stop` and warn
you.

The Raspberry Pi analog audio output uses both PWM channels. You can't
use hardware PWM and audio output at the same time.

Although it is possible to create software PWM, this isn't advised due
to latency of linux and R. `rpigpior` does not supply any software PWM
channels.

## Enabling hardware PWM on the Raspberry Pi

*Raspberry Pi OS* uses the device tree to enable hardware PWM. You'll
need to do a bit of behind-the-scenes editing to set this up.

To do this, edit /boot/config.txt as root.

1.  On the Raspberry Pi, open a terminal window

2.  Use the command `sudo nano /boot/config.txt`. This will open the
    nano text editor and display the contents of /boot/config.txt.

3.  Add the `dtoverlay` command at the end of the file. The easiest way
    to determine the command to add to `/boot/config.txt` is to ask
    `rpi_pwm()` for help. Set `pwm_debug=TRUE` then select the pins you
    want to use. If they aren't enabled, `rpi_pwm()` will stop and
    provide you with the correct `dtoverlay` string. This looks
    something like `dtoverlay=pwm,pin18,func-2` but will change
    depending on the pins you have chosen.\
    \
    If you want to do the math yourself, there is an excellent guide
    available at
    [github.com/dotnet](https://github.com/dotnet/iot/blob/main/Documentation/raspi-pwm.md).
    There is also linux documentation that discusses [Enabling one PWM
    channel](https://github.com/raspberrypi/linux/blob/04c8e47067d4873c584395e5cb260b4f170a99ea/arch/arm/boot/dts/overlays/README#L925)
    or [Enabling a dual pwm
    channel](https://github.com/raspberrypi/linux/blob/04c8e47067d4873c584395e5cb260b4f170a99ea/arch/arm/boot/dts/overlays/README#L944)

4.  Use *control-o* to write the file, then *control-x* to exit `nano`

5.  Reboot your Raspberry Pi to enact this change.

Note that if you rebuild your SD card or reinstall Raspberry Pi OS on
the OS card, you'll need to repeat this step.

## How to use `rpi_pwm()`

Here's an example:

```{r, eval=FALSE}
rpi_pwm(c(12,33), pwm_period = 50000, pwm_dutycycle = 25000)

# This will enable PWM0 and PWM1 with a period (frequency) of 50000 and a duty cycle of 25000. That is to say, 50% power.
```

Or...

```{r, eval=FALSE}
rpi_pwm(12, pwm_period = 50000, pwm_dutycycle = 10000)

# This will turn PWM0 down to 20%
```

## Handling Errors:

If you set `pwm_debug = TRUE` then `rpi_pwm()` will run diagnostics as
it goes.

```{r, eval=FALSE}
rpi_pwm(12, pwm_period = 50000, pwm_dutycycle = 10000, pwm_debug = TRUE)
```

### Invalid PWM pin:

You've called `rpi_pwm(pin_number)` where the pin (or pins) specified
don't supply hardware Pulse Width Modulation. On the Raspberry Pi, this
is only board pins 12, 32, 33, or 35. Keep in mind `rpi_pwm()` only
specifies pins as board pins - not GPIO or BCM pins. Lots of
documentation will refer to these other pin numbering schemes.

### Invalid PWM pin combination:

As you know by now, the Raspberry Pi only has two hardware PWM channels,
but each channel is connected to two pins:

-   PWM0 appears on board pins 12 (GPIO18) and 32 (GPIO12).

-   PWM1 appears on board pins 33 (GPIO13) and 35 (GPIO19).

So...if you are going to use two PWM channels, you need to use pins
connected to different channels. Since board pin 12 and 32 are both
connected to PWM0, it would be nonsense to use
`rpi_pwm(pin_number = c(12,32)`. Instead, these are the valid
combinations:

-   `rpi_pwm(pin_number = c(12,33)`
-   `rpi_pwm(pin_number = c(12,35)`
-   `rpi_pwm(pin_number = c(32,33)`
-   `rpi_pwm(pin_number = c(32,35)`

If you try to select an invalid combination and `pwm_debut = TRUE` you
will receive an error.

### PWM not enabled:

The PWM channel (or channels) you have selected are not yet enabled.
Enabling PWM is described in the above section titled [Enabling hardware
PWM on the Raspberry Pi] . The error message from `rpi_pwm` supplies the
string you'll need to paste into `/boot/config.txt`.

## Notes:

The Linux driver implementer's API guide has a chapter on PWM :
<https://www.kernel.org/doc/html/v5.10/driver-api/pwm.html>
