---
title: "rpi_i2c"
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rpigpior)
```

# About I2C on Raspberry Pi

Inter-Integrated Circuit Communication

Uses Serial Data Line (SDA, board pin 3) and Serial Clock Line (SCL,
board pin 5)

RPI has two hardware i2c buses. Software buses can also be
[created](https://raspberrypi.stackexchange.com/questions/101951/using-normal-gpio-pins-as-i2c-required)

# Installing I2C

Requires turning on i2c from Raspi-config

use `i2cdetect -y 1` to show devices connected to i2c bus

# i2c-tools

i2c-tools is already installed with Raspberry Pi Operating System

-   i2cset set one byte

-   i2cget get one byte

-   i2cdump

-   i2cdetect -y 1

-   i2ctransfer send/receive multiple bytes

<https://github.com/epsylon3/android_external_i2c-tools/tree/gingerbread/tools>

## Handling Errors:

stop("You must specify a chip_address as an integer between 0x08 and
0x77")

stop("You must specify a data_address as an integer between 0x00 and
0xFF")

-   If the i2cdetect command results in an error you have either not
    installed i2c-tools or you need to use 0 rather than 1

-   If your device is not detected and you don't see the address you
    have either not connected the device properly or have not correctly
    enabled the i2c interface
